### 2012-05-26 latex/montaje.tex
anexo de montaje

\documentclass[main]{subfiles}
\begin{document}

\chapter{Montaje}
\label{chap:camara}

Doble fueye\\
foto amortiguador\\
mostrar ruidos\\
montaje gps\\
montaje wifi?\\
montaje imu, en el centro del quad\\
placa de conversores lógicos\\

% \begin{wrapfigure}{r}{0.55\textwidth}
\begin{figure}[h!]
	\centering
		\includegraphics[width=0.7\textwidth]{./pics_montaje/beagle.jpg}
	\caption{Montaje del $BeagleBoard$}
	\label{fig:beagle}
\end{figure}
%\end{wrapfigure}

\begin{wrapfigure}{r}{0.55\textwidth}
	\begin{center}
		\includegraphics[width=0.5\textwidth]{./pics_montaje/horizontalidad.jpg}
	\end{center}
	\caption{Horizontalización}
	\label{fig:horizontalidad}
\end{wrapfigure}

\end{document}