### 2012-07-02 latex/dedicatoria.tex

\documentclass[main]{subfiles}
\begin{document}

\chapter*{}
\pagenumbering{roman} % para comenzar la numeracion de paginas en numeros romanos
\begin{flushright}
\textit{
Dedicado al DNS \\
que lo mira por TV}
\end{flushright}
\end{document}